Short-term hourly forecasting for care services

### 2021-06-22 paper/paper.Rmd
Added bits to paper

An accurate demand forecasting is crucial in Accident & Emergency and Ambulance Services in particular to depict various courses of action that can result in massive savings in terms of patient lives. 
Unability to match the staff with the demand results in patient overcrowding the in urgent and emergency care system which is a serious problem that causes challenging situations on patient flow. Also, it is related with increasing length of stay, low patient satisfaction, number of patients left the A&E without seen by staff, unexpected return visits to A&Es , increasing health care costs, inaccuracy in electronic medical record, and reported waiting times.without incurring last-minute expenses, such as overtime or supplemental staffing.

An accurate forecasting of the patient demand by hour of the day enables managers to match staff to meet anticipated patients, reconfigure units and redeploy staff. This will have many advantages for both patients, staff and the quality of healthcare services provided.

Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day. This involves the short-term decision making related to the execution of the delivery process for various health care services such as Ambulatory ,Emergency, Surgical, Inpatient, Home and Residential.

@@

limitations in current literature?

@@

1. We develop a novel methodology to forecast hourly time series for healthcare services combining a distribution fit to each hour of day separately and then adjust it using variables accounting for i) trend, ii) autocorrelation iii) temporal factors such as day of week, ... iv) special events such as public holidays, festive days and rugby and v) weather data.
2. We provide probabilistic forecasts ...
3. We benchmark the forecasting performance of the proposed methodology against regression, prophet and FASSTER using ....

@@

# Proposed model {#model}

@@

@@

## TBATS

>>>>

An accurate demand forecasting is crucial in Accident & Emergency and Ambulance Services in particular to depict various courses of action that can result in massive savings in terms of patient lives. 
Unability to match the staff with the demand results in patient overcrowding the in urgent and emergency care system which is a serious problem that causes challenging situations on patient flow. Also, it is related with increasing length of stay, low patient satisfaction, number of patients left the A&E without seen by staff, unexpected return visits to A&Es , increasing health care costs, inaccuracy in electronic medical record, and reported waiting times without incurring last-minute expenses, such as overtime or supplemental staffing.

An accurate forecast of the patient demand by hour of the day enables managers to match staff to meet anticipated patients, reconfigure units and redeploy staff. This will have many advantages for both patients, staff and the quality of healthcare services provided.

Furthermore, because the consequences of imperfact staffing are asymmetric probabilistic forecasts are necessary to make economic decicions. This asymmetry arrises because it is preferable to incure a small opportunity cost associated unerutilised staff rather than lower service levels if staff levels are insufficient.

Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day. This involves the short-term decision making related to the execution of the delivery process for various health care services such as Ambulatory, Emergency, Surgical, Inpatient, Home and Residential.

@@

limitations in current literature?
- Typically focus on daily data, hiding information about peak arrival rates. Hourly data has additional challenge of multiple seasonality.
- Deterministic, are there any examples of probabilistic?

Other papers to review and possibly cite:
- Daily visits using Prophet 10.1016/j.annemergmed.2019.08.149
- Daily using internet data 10.1016/j.annemergmed.2014.10.008
- Review paper: 10.1136/emj.2008.062380
- Others: 10.1136/emj.2010.103531, 10.1136/emj.2007.051656
- Impact of weather 10.1136/emj.2010.091058, 10.1136/emj.20.2.204

@@

1. We develop a novel methodology to forecast hourly time series for healthcare services combining a distribution fit to each hour of day separately and then adjust it using variables accounting for i) trend, ii) autocorrelation iii) temporal factors such as day of week, ... iv) special events such as public holidays, festive days and rugby and v) weather data.
2. We provide probabilistic forecasts ...
3. We benchmark the forecasting performance of the proposed methodology against regression, prophet and FASSTER using ....
4. We provide data and code enabling reproduction and refinement of the proposed approach

@@

# Proposed models {#model}

## Exponential Smoothing

## GAMLSS

@@

### Naive/Climatology

Emprical distribution for same hour-of-day and day-of-week as target time.

@@

### TBATS
